HO unification from object language to meta language

### 2024-02-15 main.tex
update tex

\newcommand{\library}[1]{\textit{#1}\xspace}
\def\stdpp{\library{stdpp}}
\def\iris{\library{iris}}
\tableofcontents

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Context
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% embed tc in ml
The motivating example for this article is our investigation for an alternative
type-class solver for \coq in \elpi\todo{citation to elpi} through the \coqelpi
plugin. This plugin gives a deep embedding of \coq terms into \elpi syntax, but
challenges arise whe attempting to solve type-class goals represented into their
corresponding \elpi representation.

\begin{itemize}
  \item about meta programming (what it is, when to use it)
  \item theorem provers (proof automation)
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Example
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\newcommand{\library}[1]{\textit{#1}\xspace}
\def\stdpp{\library{stdpp}}
\def\iris{\library{iris}}
% \tableofcontents

@@

@@

% tc motivates this paper
The motivating example for this article is our investigation for an alternative
type-class solver for \coq in \elpi\todo{citation to elpi} through the \coqelpi
plugin. This plugin gives a deep embedding of \coq terms into \elpi syntax, but
challenges arise whe attempting to solve type-class goals represented into their
corresponding \elpi representation.

% syntax of coq terms in elpi
In particular, \coq terms are encoded following the following (very simplified)
data type:

\begin{elpicode}
  kind term type.
  type const gref -> term.
  type app list term -> term.
  type fun (term -> term) -> term.
  type prod (term -> term) -> term.
\end{elpicode}

Where a \coq constant is a gref\footnote{a gref is the opaque type for \coq
identifier in \elpi} inside the \elpiIn{const} constructor, the \coq application
is embedded in the \elpiIn{app} node, having the head and the arguments of the
application expressed as a list of terms. Lambda abstractions are translated
into the node \elpiIn{fun} binding a term to another term. Finally forall
quantification on the for \coqIn{~$\forall$~x,F} into \elpiIn{forall x\ F}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Examples
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Let's take the following example:
\begin{tcolorboxSimple}
  \begin{tabular}{c|c}
    \begin{minipage}[c]{0.50\linewidth}
      \begin{coqcode}
        Instance forall_dec A (P: A ~$\to$~ Prop):
          Finite A ~$\to$~ 
          ~$\forall$~x, Decision (P x) ~$\to$~
          Decision (~$\forall$~x, P x).
      \end{coqcode}
    \end{minipage}
    &
    \begin{minipage}[c]{0.46\linewidth}
      \begin{elpicode}
        tc-Decision (prod x\ app[P, x]) :- 
          % premise for Finite A
          % premise for $\forall$x, Decision (P x)
      \end{elpicode}
    \end{minipage}
  \end{tabular}
\end{tcolorboxSimple}

The \coq instance on the left is compiled in the \elpi rule on the right. If 
we try to solve the following query (already in the \elpi syntax)

\elpiIn{tc-Decision (prod x\ app[f, y, x])}

where \coqIn{f : A -> B -> Type} and \coqIn{y: A}, we can notice that the \elpi
rule above cannot be applied; the terms \elpiIn{prod x\ app[P, x]} and
\elpiIn{prod x\ app[f, y, x]} cannot unify since the lists under the
\elpiIn{app} constructors are of different length.\todo{NOTE: the goal unify
with forall\_dec in coq xké fa unif FO approx}
